---
title: "R Notebook"
output: html_notebook
---

```{r}
library(Seurat)
library(dplyr)
library(ggplot2)
```

```{r}
outdir <- "/path/to/outdir/"

params <- read.delim("/path/to/plotting_params.txt", row.names = 1) #plotting parameters

# load the Seurat objects for the DE analysis
BCC <- readRDS("/path/to/Visium_BCC_inte.rds")
SCC <- readRDS("/path/to/Visium_SCC_inte.rds")
Mela <- readRDS("/path/to/Visium_Mela_inte.rds")
```

```{r}
# load the LR scores from stLearn run (all scores, not just significant)
BCC_B18 <- read.csv("/path/to/B18BCC_LR_score.csv", row.names = 1)
BCC_E15 <- read.csv("/path/to/E15BCC_LR_score.csv", row.names = 1)
BCC_F21 <- read.csv("/path/to/F21BCC_LR_score.csv", row.names = 1)

SCC_B18 <- read.csv("/path/to/B18SCC_LR_score.csv", row.names = 1)
SCC_E15 <- read.csv("/path/to/E15SCC_LR_score.csv", row.names = 1)
SCC_F21 <- read.csv("/path/to/F21SCC_LR_score.csv", row.names = 1)
SCC_P13 <- read.csv("/path/to/P13SCC_LR_score.csv", row.names = 1)
SCC_P30 <- read.csv("/path/to/P30SCC_LR_score.csv", row.names = 1)

mel_6747 <- read.csv("/path/to/Mela6747_LR_score.csv", row.names = 1)
mel_21031 <- read.csv("/path/to/Mela21031_LR_score.csv", row.names = 1)
mel_48974 <- read.csv("/path/to/Mela48974_LR_score.csv", row.names = 1)
mel_66487 <- read.csv("/path/to/Mela66487_LR_score.csv", row.names = 1)
```

# Set up a function to plot the samples

```{r}
DoFeaturePlot <- function(myfeature) {
  SpatialFeaturePlot(SCC, features = myfeature, images = "slice1", pt.size.factor = params["SCC_B18","dotsize"]) +
  theme(aspect.ratio = params["SCC_B18","ratio"])
  ggsave(filename = paste0(outdir, myfeature, "_", "SCC_B18", ".jpeg"))
  
SpatialFeaturePlot(SCC, features = myfeature, images = "slice1.1", pt.size.factor = params["SCC_E15","dotsize"]) +
  theme(aspect.ratio = params["SCC_E15","ratio"])
ggsave(filename = paste0(outdir, myfeature, "_", "SCC_E15", ".jpeg"))
SpatialFeaturePlot(SCC, features = myfeature, images = "slice1.2", pt.size.factor = params["SCC_F21","dotsize"]) +
  theme(aspect.ratio = params["SCC_F21","ratio"])
ggsave(filename = paste0(outdir, myfeature, "_", "SCC_F21", ".jpeg"))
SpatialFeaturePlot(SCC, features = myfeature, images = "slice1.3", pt.size.factor = params["SCC_P13","dotsize"]) +
  theme(aspect.ratio = params["SCC_P13","ratio"])
ggsave(filename = paste0(outdir, myfeature, "_", "SCC_P13", ".jpeg"))
SpatialFeaturePlot(SCC, features = myfeature, images = "slice1.4", pt.size.factor = params["SCC_P30","dotsize"]) +
  theme(aspect.ratio = params["SCC_P30","ratio"])
ggsave(filename = paste0(outdir, myfeature, "_", "SCC_P30", ".jpeg"))

SpatialFeaturePlot(BCC, features = myfeature, images = "slice1", pt.size.factor = params["BCC_B18","dotsize"]) + 
  theme(aspect.ratio = params["BCC_B18","ratio"])
ggsave(filename = paste0(outdir, myfeature, "_", "BCC_B18", ".jpeg"))
SpatialFeaturePlot(BCC, features = myfeature, images = "slice1.1", pt.size.factor = params["BCC_E15","dotsize"]) +
  theme(aspect.ratio = params["BCC_E15","ratio"])
ggsave(filename = paste0(outdir, myfeature, "_", "BCC_E15", ".jpeg"))
SpatialFeaturePlot(BCC, features = myfeature, images = "slice1.2", pt.size.factor = params["BCC_F21","dotsize"]) +
  theme(aspect.ratio = params["BCC_F21","ratio"])
ggsave(filename = paste0(outdir, myfeature, "_", "BCC_F21", ".jpeg"))

SpatialFeaturePlot(Mela, features = myfeature, images = params["mel_48974","slice"], pt.size.factor = params["mel_48974","dotsize"]) +
  theme(aspect.ratio = params["mel_48974","ratio"])
ggsave(filename = paste0(outdir, myfeature, "_", "mel_48974", ".jpeg"))
SpatialFeaturePlot(Mela, features = myfeature, images = "slice1.1", pt.size.factor = params["mel_66487","dotsize"]) +
  theme(aspect.ratio = params["mel_66487","ratio"])
ggsave(filename = paste0(outdir, myfeature, "_", "mel_66487", ".jpeg"))
SpatialFeaturePlot(Mela, features = myfeature, images = "slice1.2", pt.size.factor = params["mel_21031","dotsize"]) +
  theme(aspect.ratio = params["mel_21031","ratio"])
ggsave(filename = paste0(outdir, myfeature, "_", "mel_21031", ".jpeg"))
SpatialFeaturePlot(Mela, features = myfeature, images = "slice1.3", pt.size.factor = params["mel_6747","dotsize"]) + 
  theme(aspect.ratio = params["mel_6747","ratio"])
ggsave(filename = paste0(outdir, myfeature, "_", "mel_6747", ".jpeg"))
}
```


```{r}
# now we just want to extract the CSF1R/IL34 one
BCC_B18 <- BCC_B18[,"IL34_CSF1R", drop = FALSE]
#BCC_E15 <- BCC_E15[,"IL34_CSF1R", drop = FALSE]
BCC_F21 <- BCC_F21[,"IL34_CSF1R", drop = FALSE]

SCC_B18 <- SCC_B18[,"IL34_CSF1R", drop = FALSE]
#SCC_E15 <- SCC_E15[,"IL34_CSF1R", drop = FALSE]
#SCC_F21 <- SCC_F21[,"IL34_CSF1R", drop = FALSE]
SCC_P13 <- SCC_P13[,"IL34_CSF1R", drop = FALSE]
SCC_P30 <- SCC_P30[,"IL34_CSF1R", drop = FALSE]

mel_6747 <- mel_6747[,"IL34_CSF1R", drop = FALSE]
mel_21031 <- mel_21031[,"IL34_CSF1R", drop = FALSE]
mel_48974 <- mel_48974[,"IL34_CSF1R", drop = FALSE]
mel_66487 <- mel_66487[,"IL34_CSF1R", drop = FALSE]
```

# Add into the Seurat object so I can plot the scores

```{r}
# update with the unique sample barcodes
## first we need to find out what the barcodes are
unique(SCC[["sample"]])
rownames(SCC_P30) <- paste0(rownames(SCC_P30), "_5")
rownames(SCC_P13) <- paste0(rownames(SCC_P13), "_4")
rownames(SCC_B18) <- paste0(rownames(SCC_B18), "_1")
allSCC <- rbind(SCC_P30, SCC_P13, SCC_B18)

# update with the unique sample barcodes
## first we need to find out what the barcodes are
unique(BCC[["sample"]])
rownames(BCC_B18) <- paste0(rownames(BCC_B18), "_1")
rownames(BCC_F21) <- paste0(rownames(BCC_F21), "_3")
allBCC <- rbind(BCC_B18, BCC_F21)

# update with the unique sample barcodes
## first we need to find out what the barcodes are
unique(Mela[["sample"]])
rownames(mel_66487) <- paste0(rownames(mel_66487), "_2")
rownames(mel_48974) <- paste0(rownames(mel_48974), "_1")
rownames(mel_6747) <- paste0(rownames(mel_6747), "_4")
rownames(mel_21031) <- paste0(rownames(mel_21031), "_3")
allmel <- rbind(mel_66487, mel_48974, mel_6747, mel_21031)
```

```{r}
# now add into the Seurat object
SCC <- AddMetaData(SCC, allSCC)
BCC <- AddMetaData(BCC, allBCC)
Mela <- AddMetaData(Mela, allmel)

# get rid of the NAs
SCC_meta <- SCC[["IL34_CSF1R"]]
BCC_meta <- BCC[["IL34_CSF1R"]]
mel_meta <- Mela[["IL34_CSF1R"]]

# fix the NAs
SCC_meta[is.na(SCC_meta)] <- 0
BCC_meta[is.na(BCC_meta)] <- 0
mel_meta[is.na(mel_meta)] <- 0

# add back into metadata
SCC <- AddMetaData(SCC, SCC_meta)
BCC <- AddMetaData(BCC, BCC_meta)
Mela <- AddMetaData(Mela, mel_meta)
```

# Plot the LR scores in each sample

```{r}
DoFeaturePlot("IL34_CSF1R")
```
# Explore thresholds

Options include:
* Spots above 25% of the max score (across cancer types, or within cancer types)
* Spots above 50% of the max score (across cancer types, or within cancer types)
* Non-zero spots

```{r}
# test binary spots
SCC_meta_bin <- SCC_meta
BCC_meta_bin <- BCC_meta
Mela_meta_bin <- mel_meta

SCC_meta_bin[SCC_meta_bin > 0] <- 1
colnames(SCC_meta_bin) <- "IL34_CSF1R_binary"
BCC_meta_bin[BCC_meta_bin > 0] <- 1
colnames(BCC_meta_bin) <- "IL34_CSF1R_binary"
Mela_meta_bin[Mela_meta_bin > 0] <- 1
colnames(Mela_meta_bin) <- "IL34_CSF1R_binary"

# add back into metadata
SCC <- AddMetaData(SCC, SCC_meta_bin)
BCC <- AddMetaData(BCC, BCC_meta_bin)
Mela <- AddMetaData(Mela, Mela_meta_bin)

# Plot
DoFeaturePlot("IL34_CSF1R_binary")
```
The distribution/sparseness really varies across tissues

```{r}
# test 25% of max
maxval <- max(c(SCC_meta$IL34_CSF1R, BCC_meta$IL34_CSF1R, mel_meta$IL34_CSF1R))
threshold <- maxval/4

# make a copy
SCC_meta_25 <- SCC_meta
BCC_meta_25 <- BCC_meta
Mela_meta_25 <- mel_meta

# renumber
SCC_meta_25[SCC_meta_25 < threshold] <- 0
SCC_meta_25[SCC_meta_25 >= threshold] <- 1
colnames(SCC_meta_25) <- "IL34_CSF1R_25"

BCC_meta_25[BCC_meta_25 < threshold] <- 0
BCC_meta_25[BCC_meta_25 >= threshold] <- 1
colnames(BCC_meta_25) <- "IL34_CSF1R_25"

Mela_meta_25[Mela_meta_25 < threshold] <- 0
Mela_meta_25[Mela_meta_25 >= threshold] <- 1
colnames(Mela_meta_25) <- "IL34_CSF1R_25"

# add back into metadata
SCC <- AddMetaData(SCC, SCC_meta_25)
BCC <- AddMetaData(BCC, BCC_meta_25)
Mela <- AddMetaData(Mela, Mela_meta_25)

# Plot
DoFeaturePlot("IL34_CSF1R_25")
```

```{r}
# test 50% of max
maxval <- max(c(SCC_meta$IL34_CSF1R, BCC_meta$IL34_CSF1R, mel_meta$IL34_CSF1R))
threshold <- maxval/4

# make a copy
SCC_meta_50 <- SCC_meta
BCC_meta_50 <- BCC_meta
Mela_meta_50 <- mel_meta

# renumber
SCC_meta_50[SCC_meta_50 < threshold] <- 0
SCC_meta_50[SCC_meta_50 >= threshold] <- 1
colnames(SCC_meta_50) <- "IL34_CSF1R_50"

BCC_meta_50[BCC_meta_50 < threshold] <- 0
BCC_meta_50[BCC_meta_50 >= threshold] <- 1
colnames(BCC_meta_50) <- "IL34_CSF1R_50"

Mela_meta_50[Mela_meta_50 < threshold] <- 0
Mela_meta_50[Mela_meta_50 >= threshold] <- 1
colnames(Mela_meta_50) <- "IL34_CSF1R_50"

# add back into metadata
SCC <- AddMetaData(SCC, SCC_meta_50)
BCC <- AddMetaData(BCC, BCC_meta_50)
Mela <- AddMetaData(Mela, Mela_meta_50)

# Plot
DoFeaturePlot("IL34_CSF1R_50")
```

# How many spots do we have with each option

Even 25% looks pretty sparse in some samples

```{r}
print("binary")
table(SCC$IL34_CSF1R_binary)
table(BCC$IL34_CSF1R_binary)
table(Mela$IL34_CSF1R_binary)
print("25%")
table(SCC$IL34_CSF1R_25)
table(BCC$IL34_CSF1R_25)
table(Mela$IL34_CSF1R_25)
print("50%")
table(SCC$IL34_CSF1R_50)
table(BCC$IL34_CSF1R_50)
table(Mela$IL34_CSF1R_50)
```
# Let's test the 25% of the max (where the same max is used for all tissues)

```{r}
#want to filter only to spots expressing CSF1R as well
SCC_CSF1R <- as.data.frame(SCC@assays$Spatial@counts["CSF1R",])
SCC_CSF1R[SCC_CSF1R > 0] <- 1

BCC_CSF1R <- as.data.frame(BCC@assays$Spatial@counts["CSF1R",])
BCC_CSF1R[BCC_CSF1R > 0] <- 1

mel_CSF1R <- as.data.frame(Mela@assays$Spatial@counts["CSF1R",])
mel_CSF1R[mel_CSF1R > 0] <- 1
```

```{r}
# combine the CSF1R and 25% IL34_CSF1R info together
SCC_meta2 <- SCC[["IL34_CSF1R_25"]]
SCC_meta2 <- cbind(SCC_meta2, SCC_CSF1R)
SCC_meta2 <- mutate(SCC_meta2, totest_25pc = ifelse(IL34_CSF1R_25 == 1 & `SCC@assays$Spatial@counts["CSF1R", ]` == 1, "yes", "no"))

BCC_meta2 <- BCC[["IL34_CSF1R_25"]]
BCC_meta2 <- cbind(BCC_meta2, BCC_CSF1R)
BCC_meta2 <- mutate(BCC_meta2, totest_25pc = ifelse(IL34_CSF1R_25 == 1 & `BCC@assays$Spatial@counts["CSF1R", ]` == 1, "yes", "no"))

Mela_meta2 <- Mela[["IL34_CSF1R_25"]]
Mela_meta2 <- cbind(Mela_meta2, mel_CSF1R)
Mela_meta2 <- mutate(Mela_meta2, totest_25pc = ifelse(IL34_CSF1R_25 == 1 & `Mela@assays$Spatial@counts["CSF1R", ]` == 1, "yes", "no"))

SCC <- AddMetaData(SCC, SCC_meta2[,"totest_25pc",drop = FALSE])
BCC <- AddMetaData(BCC, BCC_meta2[,"totest_25pc",drop = FALSE])
Mela <- AddMetaData(Mela, Mela_meta2[,"totest_25pc",drop = FALSE])
```

```{r}
# how many hits
table(SCC$totest)
table(BCC$totest)
table(Mela$totest)

#table(SCC$IL34_CSF1R_25)
#table(BCC$IL34_CSF1R_25)
#table(Mela$IL34_CSF1R_25)
```


# try doing the DE analysis between the y and n groups

```{r}
Idents(SCC) <- "IL34_CSF1R_25"
Idents(BCC) <- "IL34_CSF1R_25"
Idents(Mela) <- "IL34_CSF1R_25"

markers_SCC <- FindMarkers(SCC, ident.1 = 1, ident.2 = 0, min.pct = 0.25, logfc.threshold = 0.25) %>% filter(p_val_adj <= 0.05)
markers_BCC <- FindMarkers(BCC, ident.1 = 1, ident.2 = 0, min.pct = 0.25, logfc.threshold = 0.25) %>% filter(p_val_adj <= 0.05)
markers_mel <- FindMarkers(Mela, ident.1 = 1, ident.2 = 0, min.pct = 0.25, logfc.threshold = 0.25) %>% filter(p_val_adj <= 0.05)
```

```{r}
Idents(SCC) <- "totest"
Idents(BCC) <- "totest"
Idents(Mela) <- "totest"

markers_SCC_2 <- FindMarkers(SCC, ident.1 = "yes", ident.2 = "no", min.pct = 0.25, logfc.threshold = 0.25) %>% filter(p_val_adj <= 0.05)
markers_BCC_2 <- FindMarkers(BCC, ident.1 = "yes", ident.2 = "no", min.pct = 0.25, logfc.threshold = 0.25) %>% filter(p_val_adj <= 0.05)
markers_mel_2 <- FindMarkers(Mela, ident.1 = "yes", ident.2 = "no", min.pct = 0.25, logfc.threshold = 0.25) %>% filter(p_val_adj <= 0.05)

write.table(markers_SCC_2, file = paste0(outdir, "SCC_markers_25pc.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(markers_BCC_2, file = paste0(outdir, "BCC_markers_25pc.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(markers_mel_2, file = paste0(outdir, "mel_markers_25pc.txt"), sep = "\t", quote = FALSE, col.names = NA)
```

# Run DEG on binary presence/absence column


```{r}
# combine the CSF1R and 25% IL34_CSF1R info together
SCC_meta3 <- SCC[["IL34_CSF1R_binary"]]
SCC_meta3 <- cbind(SCC_meta3, SCC_CSF1R)
SCC_meta3 <- mutate(SCC_meta3, totest_binary = ifelse(IL34_CSF1R_binary == 1 & `SCC@assays$Spatial@counts["CSF1R", ]` == 1, "yes", "no"))

BCC_meta3 <- BCC[["IL34_CSF1R_binary"]]
BCC_meta3 <- cbind(BCC_meta3, BCC_CSF1R)
BCC_meta3 <- mutate(BCC_meta3, totest_binary = ifelse(IL34_CSF1R_binary == 1 & `BCC@assays$Spatial@counts["CSF1R", ]` == 1, "yes", "no"))

Mela_meta3 <- Mela[["IL34_CSF1R_binary"]]
Mela_meta3 <- cbind(Mela_meta3, mel_CSF1R)
Mela_meta3 <- mutate(Mela_meta3, totest_binary = ifelse(IL34_CSF1R_binary == 1 & `Mela@assays$Spatial@counts["CSF1R", ]` == 1, "yes", "no"))

SCC <- AddMetaData(SCC, SCC_meta3[,"totest_binary",drop = FALSE])
BCC <- AddMetaData(BCC, BCC_meta3[,"totest_binary",drop = FALSE])
Mela <- AddMetaData(Mela, Mela_meta3[,"totest_binary",drop = FALSE])
```

```{r}
# how many hits
table(SCC$test_binary)
table(BCC$test_binary)
table(Mela$test_binary)

#table(SCC$IL34_CSF1R_25)
#table(BCC$IL34_CSF1R_25)
#table(Mela$IL34_CSF1R_25)
```

# try doing the DE analysis between the y and n groups

```{r}
Idents(SCC) <- "test_binary"
Idents(BCC) <- "test_binary"
Idents(Mela) <- "test_binary"

markers_SCC <- FindMarkers(SCC, ident.1 = "yes", ident.2 = "no", min.pct = 0.25, logfc.threshold = 0.25) %>% filter(p_val_adj <= 0.05)
markers_BCC <- FindMarkers(BCC, ident.1 = "yes", ident.2 = "no", min.pct = 0.25, logfc.threshold = 0.25) %>% filter(p_val_adj <= 0.05)
markers_mel <- FindMarkers(Mela, ident.1 = "yes", ident.2 = "no", min.pct = 0.25, logfc.threshold = 0.25) %>% filter(p_val_adj <= 0.05)

write.table(markers_SCC, file = paste0(outdir, "SCC_markers_binaryGroups.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(markers_BCC, file = paste0(outdir, "BCC_markers_binaryGroups.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(markers_mel, file = paste0(outdir, "mel_markers_binaryGroups.txt"), sep = "\t", quote = FALSE, col.names = NA)
```

# merge the files together

```{r}
# binary
markers_SCC$cancer <- "SCC"
markers_BCC$cancer <- "BCC"
markers_mel$cancer <- "mel"
binary_all <- rbind(markers_SCC, markers_BCC, markers_mel)
write.table(binary_all, file = paste0(outdir, "allmarkers_binaryGroups.txt"), sep = "\t", quote = FALSE, col.names = NA)
# 25%
markers_SCC_2$cancer <- "SCC"
markers_BCC_2$cancer <- "BCC"
markers_mel_2$cancer <- "mel"
binary_25pc <- rbind(markers_SCC_2, markers_BCC_2, markers_mel_2)
write.table(binary_25pc, file = paste0(outdir, "allmarkers_25pc.txt"), sep = "\t", quote = FALSE, col.names = NA)
```

```{r}
SCC_meta_xinnan <- SCC[[c("totest_25pc", "totest_binary")]]
BCC_meta_xinnan <- BCC[[c("totest_25pc", "totest_binary")]]
mel_meta_xinnan <- Mela[[c("totest_25pc", "totest_binary")]]
write.table(SCC_meta_xinnan, file = paste0(outdir, "SCC_spotsToTest.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(BCC_meta_xinnan, file = paste0(outdir, "BCC_spotsToTest.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(mel_meta_xinnan, file = paste0(outdir, "mel_spotsToTest.txt"), sep = "\t", quote = FALSE, col.names = NA)
```

