---
title: "Seurat Heatmaps"
output: html_notebook
---

/Volumes/SciDrive/Archive/SCC/1_sandbox/20210805_HealthyVSTumourAnalysis/4_PatientWise/g_coreGeneHeatmap/20210514_ComplexHeatmap_cancer.Rmd

# Load libraries

```{r message=FALSE, warning=FALSE}
library(Seurat)
library(dplyr)
library(pheatmap)
library(metacell)
library(RColorBrewer)
library(ComplexHeatmap)
library(ggplot2)
```

# User input

```{r}
df <- readRDS("/Volumes/SciDrive/Archive/SCC/0_data/SlimSCC_NoScaleRNANormAddXinnanMeta.RDS")
DefaultAssay(df) <- "RNA"
```



```{r}
outdir <- "/Volumes/SciDrive/Archive/SCC/1_sandbox/20210805_HealthyVSTumourAnalysis/4_PatientWise/g_coreGeneHeatmap/outdir/"

clusterID <- "sample_ident"

cancer_markers <- read.delim("/Volumes/SciDrive/Archive/SCC/1_sandbox/20210805_HealthyVSTumourAnalysis/4_PatientWise/d_coreGenes/outdir/all_cancerExclusive_PatientOverlapResults.txt")

#healthy_markers <- read.delim("/Volumes/SciDrive/Archive/SCC/1_sandbox/20210805_HealthyVSTumourAnalysis/4_PatientWise/d_coreGenes/outdir/all_normalExclusive_PatientOverlapResults.txt")

desiredClusterOrder <- c("P1_N", "P2_N", "P3_N", "P4_N", "P5_N",
                         "P1_SCC", "P2_SCC1", "P2_SCC2", "P3_IEC", "P4_SCC", "P5_SCC_BCC")

sampleID <- "cancer"
```

```{r}

ann_colors = list(
    status = c(`Cancer` = "deeppink3",
              `Normal` = "turquoise3"),
    patient = c(
  "P1_N" = "tomato4", 
  "P1_SCC" = "tomato1",
  "P2_N" = "darkgoldenrod4",
  "P2_SCC1" = "gold1",
  "P2_SCC2" = "lightgoldenrod1",
  "P3_N" = "olivedrab4",
  "P3_IEC" = "olivedrab2",
  "P4_N" = "cadetblue4",
  "P4_SCC" = "cadetblue2",
  "P5_N" = "lightpink4",
  "P5_SCC_BCC" = "lightpink1")
    )
```

# Get everything set up correctly

```{r}
Idents(df) <- clusterID
cancer <- as.vector(cancer_markers$genes)
#healthy <- as.vector(healthy_markers$genes)
```

# Scale the data, and make a count matrix for the heatmap

After running my Seurat pipeline, there is no scaledata slot, so I need to re-run scaledata for RNA, as per: https://github.com/satijalab/seurat/issues/2960
Andrew Butler: "Here the DoHeatmap function is trying to pull values from the scale.data slot of the RNA assay. However, it doesn't look like you ran ScaleData on that assay and thus the slot is empty. If you want to plot a heatmap of the scaled RNA data, you only need to run ScaleData before making that DoHeatmap call (not the other functions you list)."

Normally when scaling the data, you'd either scale all genes or just the 2000 variable genes. Here, we only scale the overlap markers to save time and make sure all our genes of interest get scaled

```{r}
DefaultAssay(df) <- "RNA"
df <- ScaleData(df, features = cancer)
```

```{r}
countdata_v2 <- df@assays$RNA@scale.data[which(rownames(df@assays$RNA@scale.data) %in% cancer),]
cancer2 <- cancer[cancer %in% rownames(countdata_v2)]
countdata_v2 <- countdata_v2[cancer2,] #reorder to match the plot made by 
# reorder columns based on the cell type, and thus re-order genes too so it looks nice
# order the cells by category
myorder <- as.data.frame(Idents(df))
colnames(myorder) <- c(clusterID)
myorder$sample_ident <- factor(myorder$sample_ident, levels=desiredClusterOrder)
myorder <- myorder[order(myorder$sample_ident),,drop = FALSE]

countdata_v3 <- countdata_v2[cancer2,rownames(myorder)]
```

# Make annotations

```{r}
annotations <- df[[c(clusterID, "orig_ident")]]
annotations$patient <- annotations$sample_ident
annotations$patient <- gsub("_N", "", annotations$patient)
annotations$patient <- gsub("_SCC_BCC", "", annotations$patient)
annotations$patient <- gsub("_SCC[1-9]", "", annotations$patient)
annotations$patient <- gsub("_SCC", "", annotations$patient)
annotations$patient <- gsub("_IEC", "", annotations$patient)

# make sure the annotations are in the right order
annotations <- annotations[rownames(myorder),]
colnames(annotations) <- c("sample_ident", "orig_ident", "patient")
annotations$orig_ident <- factor(annotations$orig_ident, levels=c("Normal", "Cancer"))

# specify ComplexHeatmap annotation objects
ann_clust = HeatmapAnnotation(status = annotations$orig_ident,
                              patient = annotations$sample_ident,
                              col = list(status = ann_colors$status,
                                         patient = ann_colors$patient),
                              #simple_anno_size = unit(1, "cm"), #height of annotation bar
                              show_legend = FALSE,
                              show_annotation_name = FALSE)

# ann_time = HeatmapAnnotation(time = annotations$time,
#                              col = list(time = ann_colors$time),
#                              #simple_anno_size = unit(1, "cm"), #height of annotation bar
#                              show_legend = TRUE,
#                              show_annotation_name = FALSE,
#                              annotation_legend_param = list(title = NULL,
#                                                             at = c("d1", "d3", "d7"),
#                                                             labels = c("day 1", "day 3", "day 7"),
#                                                             legend_direction = "horizontal", nrow=1))
```

# plot a heatmap to help choose the colour scale

```{r}
hist(countdata_v3)
```

```{r}
CellPalette = circlize::colorRamp2(c(min(countdata_v3), -2, -1, 0, 1, 2, max(countdata_v3)), c("darkblue", "blue", "lightskyblue", "white", "lightpink", "red", "darkred"))

ht_allgenes <- Heatmap(countdata_v3,
        cluster_rows = FALSE, cluster_columns = FALSE,
        show_column_names = FALSE,
       col = CellPalette, 
        row_names_gp = gpar(fontsize = 6),
        row_order = cancer, column_order = rownames(myorder),
        top_annotation = ann_clust,
        # legend
         heatmap_legend_param = list(title = NULL,
                                     labels_gp = gpar(fontsize = 6),
                                     direction = "horizontal", nrow=1),
        # split up clusters (cols) and additional genes (rows)
        column_split = annotations$orig_ident, column_title = NULL,
        column_gap = unit(0.5, "mm"),
        use_raster = TRUE, raster_quality = 1,
        width = unit(5, "cm") #based on current figure
        )

pdf(paste0(outdir, sampleID, "_heatmap.pdf"), height = 6)
draw(ht_allgenes, annotation_legend_side = "bottom", heatmap_legend_side = "bottom")
dev.off()

pdf(paste0(outdir, sampleID, "_heatmap_med.pdf"), height = 4)
draw(ht_allgenes, annotation_legend_side = "bottom", heatmap_legend_side = "bottom")
dev.off()

pdf(paste0(outdir, sampleID, "_heatmap_sml.pdf"), height = 3.5)
draw(ht_allgenes, annotation_legend_side = "bottom", heatmap_legend_side = "bottom")
dev.off()
```

# Now plot for only the genes in all 5 patients

```{r}
countdata_v3_5patients <- countdata_v3[rownames(countdata_v3) %in% cancer_markers$genes[cancer_markers$nPatients == 5],]
cancer_5patients <- cancer_markers[cancer_markers$nPatients == 5,"genes"]
hist(countdata_v3_5patients)
```

```{r}
ht_5ptgenes <- Heatmap(countdata_v3_5patients,
        cluster_rows = FALSE, cluster_columns = FALSE,
        show_column_names = FALSE,
       col = CellPalette, 
        row_names_gp = gpar(fontsize = 6),
        row_order = cancer_5patients, column_order = rownames(myorder),
        top_annotation = ann_clust,
        # legend
         heatmap_legend_param = list(title = NULL,
                                     labels_gp = gpar(fontsize = 6),
                                     direction = "horizontal", nrow=1),
        # split up clusters (cols) and additional genes (rows)
        column_split = annotations$orig_ident, column_title = NULL,
        column_gap = unit(0.5, "mm"),
        use_raster = TRUE, raster_quality = 1,
        width = unit(5, "cm") #based on current figure
        )

pdf(paste0(outdir, sampleID, "5patientgenes_heatmap.pdf"), height = 6)
draw(ht_5ptgenes, annotation_legend_side = "bottom", heatmap_legend_side = "bottom")
dev.off()

pdf(paste0(outdir, sampleID, "5patientgenes_heatmap_med.pdf"), height = 4)
draw(ht_5ptgenes, annotation_legend_side = "bottom", heatmap_legend_side = "bottom")
dev.off()

pdf(paste0(outdir, sampleID, "5patientgenes_heatmap_sml.pdf"), height = 3.5)
draw(ht_5ptgenes, annotation_legend_side = "bottom", heatmap_legend_side = "bottom")
dev.off()
```


# then plot for (1) only the 5-patient genes and (2) the heatlyh genes
# save as 3.5 inches high to fit in figure

# save the legend as a separate file

```{r}
lgnd = Legend(names(ann_colors$status), legend_gp = gpar(fill = ann_colors$status))
pdf(paste0(outdir, "legend_status.pdf"))
draw(lgnd)
dev.off()

lgnd = Legend(names(ann_colors$patient), legend_gp = gpar(fill = ann_colors$patient))
pdf(paste0(outdir, "legend_patient.pdf"))
draw(lgnd)
dev.off()
```


